---
title: "Class 11 HW"
author: "Marwa Mohammad Azim"
format: gfm
---
#Section 4: Population Scale Analysis

One sample is obviously not enough to know what is happening in a population. You are
interested in assessing genetic differences on a population scale. So, you processed about
~230 samples and did the normalization on a genome level. Now, you want to find whether
there is any association of the 4 asthma-associated SNPs (rs8067378...) on ORMDL3
expression.

How many samples do we have?

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

```{r}
nrow(expr)
```


```{r}
table(expr$geno)
```

>Q13: Read this file into R and determine the sample size for each genotype and their
corresponding median expression levels for each of these genotypes.

The sample size for each genotype is as follows:
A/A 108
A/G 233
G/G 121

The corresponding median expression levels for each of these genotypes are:

A/G	28.96038	
A/G	20.24449	
A/A	31.32628	
A/A	34.11169	
G/G	18.25141	
A/A 32.89721	

```{r}
library(ggplot2)
```

Lets make a boxplot

```{r}
ggplot(expr) + aes(geno, exp, fill=geno) + 
  geom_boxplot(notch=TRUE)
```


>Q14: Generate a boxplot with a box per genotype, what could you infer from the relative
expression value between A/A and G/G displayed in this plot? Does the SNP effect the
expression of ORMDL3?

The relative expression value between A/A and G/G displayed in this plot are very much different. Having G/G in this location is definitely associated with having a reduced expression of the ORMDL3 gene. 
