---
title: "Class 10:Halloween Mini Project"
author: "Marwa"
format: gfm
---

```{r}
webshot::install_phantomjs()
```


#1.Importing candy data

```{r}
candy_file <- "candy-data.csv"

candy <- read.csv ("https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv", row.names = 1)
head(candy)
```

> Q1. How many different candy types are in this dataset?

```{r}
dim(candy)
```
There are 85 different candy types in this dataset.

> Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```
There are 38 fruity candy types in the dataset.

#2. What is your favorate candy?

```{r}
candy["Twix", ]$winpercent
```

>Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

```{r}
candy["Twix", ]$winpercent
```
My favorite candy in the dataset is Twix and it's winpercent value is 81.64291.

>Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent
```
The winpercent value for Kit Kat is 76.7686.

>Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```
The winpercent value for Tootsie Roll Snack Bars is 49.6535.

```{r}
library("skimr")
skim(candy)
```

>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

The `winpercent` variable looks to be on a different scale to the majority of the other columns in the dataset. `Winpercent` measures as a percent out of 100.


>Q7. What do you think a zero and one represent for the candy$chocolate column?

A zero means that the candy does not contain chocolate and one means that the candy does contain chocolate.

>Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```

>Q9. Is the distribution of winpercent values symmetrical?

The distribution of winpercent values are not symmetrical.

>Q10. Is the center of the distribution above or below 50%?

The center of the distribution is below 50%, it's from 35-45%.

>Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
mean(candy$winpercent[as.logical(candy$chocolate)])
```

```{r}
mean(candy$winpercent[as.logical(candy$fruity)])
```

On average chocolate is ranked higher than fruit candy.

>Q12. Is this difference statistically significant?

```{r}
t.test(candy$winpercent[as.logical(candy$chocolate)],candy$winpercent[as.logical(candy$fruity)])
```

This difference is statistically significant because the p-value is 2.871e-08.

#3. Overall Candy Rankings

>Q13. What are the five least liked candy types in this set?

```{r}
library(dplyr)
least_liked <- rownames(candy %>% arrange(winpercent) %>% head(5))
least_liked
```

The five least liked candy types in this set are 'Nik L Nip', 'Boston Baked Beans', 'Chiclets', 'Super Bubble', and 'Jawbusters'.

>Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
most_liked <- rownames(candy %>% arrange(desc(winpercent)) %>% head(5))
most_liked
```

The top 5 all time favorite candy types out of this set are 'ReeseOs Peanut Butter cup', ReeseOs Miniatures','Twix','Kit Kat' 'Snicker'.

>Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```

>Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col()
```

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```

>Q17. What is the worst ranked chocolate candy?

The worst ranked chocolate candy is Sixlets.

>Q18. What is the best ranked fruity candy?

The best ranked fruity candy is Starburst.

#4. Taking a look at pricepercent

```{r}
library(ggrepel)

ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```



>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

ReeseOs Minitaures is the highest ranked in terms of winpercent for the least money because it further along the winpercent and further down the price percent.

>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

The top 5 most expensive candy types in the dataset are 'Nik L. Nip', 'Nestle Smarties', 'Ring pop', 'HersheyÕs Krackel', 'HersheyÕs Milk Chocolate' and the least popular is Nik L Nip.


>Q21. Make a barplot again with geom_col() this time using pricepercent and then improve this step by step, first ordering the x-axis by value and finally making a so called “dot chat” or “lollipop” chart by swapping geom_col() for geom_point() + geom_segment().

```{r}
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_col()
```


```{r}
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
    geom_point()
```


#5. Exploring the correlation structure

```{r}
library(corrplot)
```

```{r}
cij <- cor(candy)
corrplot(cij)
```

>Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

'Chocolate' and 'fruity' are the two variables that are anti-correlated.

>Q23. Similarly, what two variables are most positively correlated?

'Chocolate' and 'bar' are the two variables that are most positively correlated.


#6. Principal Component Analysis


```{r}
pca <- prcomp(candy, scale = TRUE)
summary(pca)
```


```{r}
plot(pca$x[,1:2])
```


```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

```{r}
my_data <- cbind(candy, pca$x[,1:3])
```

```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```



```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```



```{r}
library(plotly)
```

```{r}
ggplotly(p)
```


```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")

```

>Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

The original variables are picked up strongly by PC1 in the positive direction are fruity, hard, and pluribus. This makes sense because most fruity candy are hard and come in a bag or box of multiple candies. These 3 variables tend to be correlated. 

